Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-20 aistats.tex
Add asymptotic derivation

\subsection{Optimal $\mu_0$ is larger than $\mu^*$ for small $n_0$ and small $n$}

The upper bound~\eqref{eq:MAP_rate} has a bias term that is $0$ when $\frac{mu_0}{\mu^*} = 1 + \inv{n_0}$, e.g. for large values of $n_0$, it is $\mu_0=\mu^*$ is the best prior, but for small $n_0$, one better sets larger values for $\mu_0$. In Figure~\ref{fig:optimal_n0}, we observe this behavior numerically.

@@

\section{ASYMPTOTIC DERIVATION}
\label{app:asymptote}

\rlp{simple stuff}

>>>>

\subsection{On the Choice of a Prior}
The optimal $\mu_0$ is larger than $\mu^*$ for small $n_0$ and small $n$.
Indeed, the upper bound~\eqref{eq:MAP_rate} has a bias term that is $0$ when $\frac{mu_0}{\mu^*} = 1 + \inv{n_0}$, e.g. for large values of $n_0$, it is $\mu_0=\mu^*$ is the best prior, but for small $n_0$, one better sets larger values for $\mu_0$. In Figure~\ref{fig:optimal_n0}, we observe this behavior numerically.

@@

\section{ASYMPTOTIC DERIVATION}
\label{app:asymptote}
In this section we fill-in the lines of \S\ref{ssec:asymptote} to prove Equation~\eqref{eq:asymptote}.
Approximating $\conj$ with a second order Taylor expansion yields
\aligns{
    \bregmanconj(\m^* ; \m)
    &= \frac{1}{2}\norm{\m - \m^*}^2_{\mF}
    + O(\norm{\m - \m^*}^3),
}
where the norm is induced by the matrix
\aligns{
    \mF
    :=\nabla^2\conj(\m^*)
    = \nabla^2\logpart(\nat^*)^{-1}
    = \Cov_{\nat^*}(T)^{-1} 
}
where the second equality is a general property of convex conjugates.
Plugging the MLE~\eqref{eq:defMLE} into this quadratic and expanding it yields
\aligns{
	\E \half \norm{\inv{n}  \sum_i T_i - \m^* }_\mF^2 
	&=\inv{2n^2} \sum_i \E \norm{T_i - \m^* }_\mF^2 
	+ \inv{2n^2} \sum_{i\neq j}\expect{T_i - \m^*}^\top \mF \overbrace{\expect{T_j - \m^*}}^{0} ]\\
	&=\inv{2n}  \E \norm{T_1 - \m^* }_\mF^2 \\
	& = \inv{2n} \Tr(\mF\ \expect{(T_1 - \m^*) (T_1 - \m^*)^\top}) \\
	&= \inv{2 n} \Tr(\mF \Cov_{\nat^*}(T)) = \frac{d}{2n} 
}
where on the first line we used independence of samples, on the second line we used the fact that samples are identically distributed, and on the third line we used the  trace trick along with the linearity of the trace $\Tr$.
On the way, this also proves that for the MLE $\norm{\m - \m^*}^3 \in O(n^{-\half[3]})$.
This yield the final rate for the MLE
\begin{align}
	\E \bregmanconj \paren{\E [T(X)] ; \inv{n}  \sum_i T_i}
	= \frac{d}{2n} + O(n^{- \frac{3}{2}} ) \; .
\end{align}
For the MAP~\eqref{eq:defMAP}, the quadratic decomposes into bias and variance
\begin{align}
	\E \half \norm{\mu^* -  \frac{n_0 \mu_0 + \smallsum_i T(x_i)}{n_0+n} }^2_{\mSigma^{-1}}
	&= \frac{n d}{2(n+n_0)^2}  +  \frac{n_0^2}{(n+n_0)^2} \half \norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} \\
	&= \frac{d}{2n} + O\left(\frac{1 + \norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} }{n^2} \right) \; .
\end{align}
This $O(n^{-2})$ term is dominated by the $O(n^{-\half[3]})$ term from the quadratic approximation of the Bregman, yielding the same first order rate as for the MLE
\begin{align}
	\E \bregmanconj \paren{\E [T(X)] ; \frac{n_0 \mu_0 + \smallsum_i T(x_i)}{n_0+n} }
	= \frac{d}{2n} + O(n^{- \frac{3}{2}} ) \; .
\end{align}
